Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-12-29 edits/barryedits.tex
more minor edits
more edits

\begin{enumerate}
\item Page 1. Column 2. I've left the sentence: ``The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.'' in.  If I remove it, Bill and Andre will want to put it back.  I believe it does give the correct intuition, even if it is imprecise.
%\item Regarding the ``conditioned upon minimisation with respect to'' phrases, I have replaced ``conditioned upon'' with ``concentrated by''.  This is perhaps a bit lazy, but I 
\item Page 6, column 2, you want me to add reference for the polynomial phase estimation papers, where we talk about differentiability of the objective function at the estimator.  I've added a reference to the article on arxiv.  We can change it to the trans. sig. proc. papers later.
\item Page 6, column 2, have left in the definition of $o_P(1)$, since it doesn't hurt.
\item Page 10 column 2, I have left ``be $N$ points uniformly spaced on the interval $[-\pi, \pi)$'' since it doesn't hurt.
\item Page 11, column 1, about the phrase ``where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$''.  I want to leave this.  The phrase is taken from one of David Pollard's papers.  It is informal, but I know that if I were personally reading this for the first time, I would be scratching my head at this point until I realised that expectation has been moved outside the supremum.
\item Page 11, column 2, about the Lipschitz continuous statement.  Obviously the term involving $\cos(\phi)$ needs only the mean value theorem.  However the term involving $\cos\fracpart{\phi}$ isn't always differentiable so the mean value theorem doesn't always apply.  It's obviouly Lipschitz continuous though, just plot it.  We can add a proof if you really think it's necessary.
\item Page 12, column 1, ``for $i \in D$ such that $R_i > 0$.'' I think this is necesary, otherwise the surrounding statement is not always true.
\item Lemma 11, I would also like there to be a nicer proof!  I haven't found one though.
\item 
\end{enumerate}

>>>>

\begin{enumerate}
\item Page 1. Column 2. I've left the sentence: ``The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.'' in.  If I remove it, Bill and Andre will want to put it back.  I believe it does give the correct intuition, even if it is imprecise.
%\item Regarding the ``conditioned upon minimisation with respect to'' phrases, I have replaced ``conditioned upon'' with ``concentrated by''.  This is perhaps a bit lazy, but I 
\item Page 6, column 2, you want me to add reference for the polynomial phase estimation papers, where we talk about differentiability of the objective function at the estimator.  I've added a reference to the article on arxiv.  We can change it to the trans. sig. proc. papers later.
\item Page 6, column 2, have left in the definition of $o_P(1)$, since it doesn't hurt.
\item Page 10 column 2, I have left ``be $N$ points uniformly spaced on the interval $[-\pi, \pi)$'' since it doesn't hurt.
\item Page 11, column 1, about the phrase ``where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$''.  I want to leave this.  The phrase is taken from one of David Pollard's papers.  It is informal, but I know that if I were personally reading this for the first time, I would be scratching my head at this point until I realised that expectation has been moved outside the supremum.
\item Page 11, column 2, about the Lipschitz continuous statement.  Obviously the term involving $\cos(\phi)$ needs only the mean value theorem.  However the term involving $\cos\fracpart{\phi}$ isn't always differentiable so the mean value theorem doesn't always apply.  It's obviously Lipschitz continuous though, just plot it.  We can add a proof if you really think it's necessary.
\item Page 12, column 1, ``for $i \in D$ such that $R_i > 0$.'' I think this is necessary, otherwise the surrounding statement is not always true.
\item Lemma 11, I would also like there to be a nicer proof!  I haven't found one though.
\item You wrote
\[
R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda) = \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) \chi(\phi,\lambda) dr d\phi
\]
in your notes.  I fixed this to
\[
R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda) = \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) e^{j\fracpart{\phi}} \chi(\phi,\lambda) dr d\phi.
\]
Everything then works as expected.
\end{enumerate}

### 2012-12-29 paper.tex
more edits

\begin{lemma}\label{lem:q1k1parts}
Put
\[
q_1(\lambda) + j k_1(\lambda) = \expect\big( R_1 e^{j(\lambda + \Phi_1)} - R_1 e^{j\Phi_1}  \big).
\]
It follows that $q_1(\lambda) = o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$
\end{lemma}
\begin{proof}
We have
\[
q_1(\lambda) + j k_1(\lambda) = (e^{j\lambda} - 1) \expect R_1e^{j\Phi_1} = e^{j\lambda} - 1
\]
since the mean of $R_1e^{j\Phi_1}$ is 1.  Thus $q_1(\lambda) = \cos(\lambda) - 1$ and $k_1(\lambda) = \sin(\lambda)$.  Now $q_1(\hat{\lambda}_L) = o_P(1)$ since $\hat{\lambda}_L$ converges almost surely to zero as $L\rightarrow\infty$.  By a Taylor expansion of $\sin(\lambda)$ about $\lambda = 0$ we obtain $k_1(\lambda) = \lambda \big( 1 + \zeta(\lambda) \big)$  where $\zeta(\lambda)$ is a continuous function with $\zeta(0) = 0$.  Thus $\zeta(\hat{\lambda}_L) = o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$.
\end{proof}

\begin{lemma}\label{lem:q2k2parts}
Put
\[
q_2(\lambda) + j k_2(\lambda) = \expect\big( R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big).
\]
It follows that $q_2(\hat{\lambda}_L) = o_P(1)$ and $k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H + o_P(1) \big)$ where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
\end{lemma}
\begin{proof}
Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $q_2(\lambda)$ and $k_2(\lambda)$ around $\lambda = 0$ that is relevant.  We will examine $q_2(\lambda)$ and $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.  An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.  To keep our notation clean put 
\[
\psi_k = \frac{2\pi}{M}k + \frac{\pi}{M}
\] 
with $k \in \ints$.  When $\Phi_1 \in [\psi_{k-1}, \psi_k - \lambda)$,
\begin{align*}
e^{j\fracpart{\lambda + \Phi_1}} - e^{j\fracpart{\Phi_1}} &=  e^{j\big(\lambda + \Phi_1 - \tfrac{2\pi}{M}k\big)} - e^{j\big( \Phi_1- \tfrac{2\pi}{M}k\big)} \\
&= \big(e^{j\lambda} - 1 \big) e^{j\big(\Phi_1- \tfrac{2\pi}{M}k\big)} \\
&= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}},
\end{align*}
and when $\Phi_1 \in [\psi_{k} - \lambda, \psi_k)$,
\begin{align*}
e^{j\fracpart{\lambda + \Phi_1}} - &e^{j\fracpart{\Phi_1}} \\
&=  e^{j\big(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}\big)} - e^{j\big( \Phi_1- \tfrac{2\pi}{M}k\big)} \\
&= \big(e^{j\big(\lambda - \tfrac{2\pi}{M}\big)} - 1 \big) e^{j\fracpart{\Phi_1}} \\
&= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}}  + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)e^{j\fracpart{\Phi_1}}.
\end{align*}
Thus, when $\Phi_1 \in [\psi_{k-1}, \psi_k)$,
\begin{align*}
e^{j\fracpart{\lambda + \Phi_1}} - e^{j\fracpart{\Phi_1}} &= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}} \\
&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)e^{j\fracpart{\Phi_1}} \chi_k(\Phi_1,\lambda)
\end{align*}
where 
\[
\chi_k(\Phi_1,\lambda) = \begin{cases}
1, & \Phi_1 \in [\psi_{k} - \lambda, \psi_k)  \\
0, & \text{otherwise}.
\end{cases}
\]
Now
\begin{align*}
q_2(\lambda) + j k_2(\lambda) &= \big(e^{j\lambda} - 1 \big) h_2(0) \\
&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big)
\end{align*}
since $\expect R_1 e^{j\fracpart{\Phi_1}} = h_2(0)$ as a result of Lemma~\ref{lem:expectImRfracpart} and where $\chi(\Phi_1, \lambda) = \sum_{k \in \ints}\chi_k(\Phi_1, \lambda)$.
Now
\begin{align*}
 \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big) &= \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) \chi(\phi,\lambda) dr d\phi \\
&= \sum_{k\in\ints}\int_{0}^{2\pi} g(\phi) \chi_k(\phi,\lambda) d\phi \\
&= \sum_{k=0}^{M-1}\int_{\psi_k-\lambda}^{\psi_k} g(\phi)  d\phi \\
&= \lambda \sum_{k=1}^{M-1} g(\phi_k) 
\end{align*}
where for each $k$, $\phi_k \in (\psi_k-\lambda, \psi_k)$.  The third line above follows from
\end{proof}

\begin{lemma}\label{lem:expectImRfracpart}
$\expect R_1 \sin\fracpart{\Phi_1} = 0$.
\end{lemma}
\begin{IEEEproof}
Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
\begin{align*}
\expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi.
\end{align*}
The proof is immediate since $g(\phi)$ is even and $\sin\fracpart{\phi}$ is odd.
% where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.  Proposition~\ref{prop:gg2cont} shows that $g$ is even on $(-\pi,\pi)$.  Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.  So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
% \[
% g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
% \]
% and
% \[
% \sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
% \]
% Now,
% \begin{align*}
% I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
% &\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
% &= 0
% \end{align*}
% for all integers $k$ because
% \begin{align*}
% \int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
% &= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
% \end{align*}
% Observe that the sets
% \[
%  [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
% \]
% partition $[-\pi, \pi]$, and so
% \[
% \expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
% \]
\end{IEEEproof}

@@

\begin{lemma}\label{lem:HLtoG} $\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1)$.
\end{lemma}
\begin{IEEEproof}
The argument is similar to that used in Lemma~\ref{lem:Qconv}.  First observe that
\begin{align*}
T_L(\lambda) &= \frac{\sabs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\sabs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
%&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
%& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
&=  \big( p + o(L^{-1/2}) \big) \expect R_1\cos(\lambda + \Phi_1) \\
&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1},
\end{align*}
and because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ (by assumption in Theorem~\ref{thm:normality}), we have
\begin{align*}
\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
\end{align*}
where
\begin{align}
q_1(\lambda) &= \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big) \;\;\; \text{and} \nonumber \\
q_2(\lambda) &= \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
\end{align}
We analyse $q_1(\lambda)$ first.  By a Taylor expansion of $\cos(\lambda + \Phi_1)$ about $\lambda = 0$,
\[
\cos(\lambda + \Phi_1) - \cos(\Phi_1) = -\lambda \big( \sin(\Phi_1) + \zeta(\lambda) \big),
\]
where $\zeta$ is a continuous function with $\zeta(0) = 0$. So,
\begin{align*}
q_1(\lambda) &= -\expect R_1\lambda(\sin(\Phi_1) +  \zeta(\lambda)) \\
&= -\lambda \big( \expect R_1\sin(\Phi_1) +  \zeta(\lambda) \expect R_1 \big) \\
&= -\lambda \zeta(\lambda) \expect R_1,
\end{align*}
since $\expect R_1\sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}.  Now, 
\begin{align*}
\sqrt{L} q_1(\hat{\lambda}_L) = -\sqrt{L} \hat{\lambda}_L \zeta(\hat{\lambda}_L) \expect R_1 = o_P(1)
\end{align*}
because $\sqrt{L} \hat{\lambda}_L$ converges in distribution and $\zeta(\hat{\lambda}_L) = o_P(1)$.  An argument analogous to that used in Lemma~\ref{lem:k2conv} shows that $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.  Thus, 
\[
\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1) = o_P(1)
\]
as required.
\end{IEEEproof}

>>>>

\begin{lemma}\label{lem:q1k1parts}
Put
\[
q_1(\lambda) + j k_1(\lambda) = \expect\big[ R_1 e^{j(\lambda + \Phi_1)} - R_1 e^{j\Phi_1}  \big].
\]
We have $q_1(\lambda) = \hat{\lambda}_L o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$.
\end{lemma}
\begin{proof}
We have
\[
q_1(\lambda) + j k_1(\lambda) = (e^{j\lambda} - 1) \expect R_1e^{j\Phi_1} = e^{j\lambda} - 1
\]
since the mean of $R_1e^{j\Phi_1}$ is 1.  By a first order expansion about $\lambda = 0$ we obtain
\[
q_1(\lambda) + j k_1(\lambda) = \lambda\big( j + O(\lambda) \big)
\]
Since $\hat{\lambda}_L$ converges almost surely to zero as $L\rightarrow\infty$ it follows that $O(\hat{\lambda}_L) = o_P(1)$.  Thus
\[
q_1(\hat{\lambda}_L) + j k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big( j + o_P(1) \big)
\]
and the lemma follow by taking real and imaginary parts.
\end{proof}

\begin{lemma}\label{lem:q2k2parts}
Put
\[
q_2(\lambda) + j k_2(\lambda) = \expect\big[ R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big].
\]
We have $q_2(\hat{\lambda}_L) = \hat{\lambda}_L o_P(1)$ and $k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H + o_P(1) \big)$ where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
\end{lemma}
\begin{proof}
Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $q_2(\lambda)$ and $k_2(\lambda)$ around $\lambda = 0$ that is relevant.  We will examine $q_2(\lambda)$ and $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.  An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.  To keep our notation clean put 
\[
\psi_k = \frac{2\pi}{M}k + \frac{\pi}{M}
\] 
with $k \in \ints$.  When $\Phi_1 \in [\psi_{k-1}, \psi_k - \lambda)$,
\begin{align*}
e^{j\fracpart{\lambda + \Phi_1}} - e^{j\fracpart{\Phi_1}} &=  e^{j\big(\lambda + \Phi_1 - \tfrac{2\pi}{M}k\big)} - e^{j\big( \Phi_1- \tfrac{2\pi}{M}k\big)} \\
&= \big(e^{j\lambda} - 1 \big) e^{j\big(\Phi_1- \tfrac{2\pi}{M}k\big)} \\
&= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}},
\end{align*}
and when $\Phi_1 \in [\psi_{k} - \lambda, \psi_k)$,
\begin{align*}
e^{j\fracpart{\lambda + \Phi_1}} - &e^{j\fracpart{\Phi_1}} \\
&=  e^{j\big(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}\big)} - e^{j\big( \Phi_1- \tfrac{2\pi}{M}k\big)} \\
&= \big(e^{j\big(\lambda - \tfrac{2\pi}{M}\big)} - 1 \big) e^{j\fracpart{\Phi_1}} \\
&= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}}  + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)e^{j\fracpart{\Phi_1}}.
\end{align*}
Thus, when $\Phi_1 \in [\psi_{k-1}, \psi_k)$,
\begin{align*}
e^{j\fracpart{\lambda + \Phi_1}} - e^{j\fracpart{\Phi_1}} &= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}} \\
&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)e^{j\fracpart{\Phi_1}} \chi_k(\Phi_1,\lambda)
\end{align*}
where 
\[
\chi_k(\Phi_1,\lambda) = \begin{cases}
1, & \Phi_1 \in [\psi_{k} - \lambda, \psi_k)  \\
0, & \text{otherwise}.
\end{cases}
\]
Now
\begin{align}
q_2(\lambda) + j k_2(\lambda) &= \expect\big[ R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big] \nonumber \\
&= \big(e^{j\lambda} - 1 \big) h_2(0)  + e^{j\lambda} B(\lambda) \label{eq:q2k2withB}
\end{align}
since $\expect R_1 e^{j\fracpart{\Phi_1}} = h_2(0) + \expect R_1 \sin\fracpart{\Phi_1} = h_2(0)$ as a result of Lemma~\ref{lem:expectImRfracpart} and where
\begin{equation}\label{eq:Blambdadefn}
B(\lambda) = \big( e^{-j\tfrac{2\pi}{M}} - 1 \big)\expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)
\end{equation}
and $\chi(\Phi_1, \lambda) = \sum_{k \in \ints}\chi_k(\Phi_1, \lambda)$.

Now $e^{j\hat{\lambda}_L} = 1 + o_P(1)$ and $e^{j\hat{\lambda}_L} - 1 = \hat{\lambda}_L \big( j  + o_P(1) \big)$ by the argument in Lemma~\ref{lem:q1k1parts} and
\[
B(\hat{\lambda}_L) = -\hat{\lambda}_L\left(2j\sin(\tfrac{\pi}{M}) \sum_{k=0}^{M-1} g(\psi_k) + o_P(1)\right) 
\]
by Lemma~\ref{lem:Blambdaconv}.  Combining these results into~\eqref{eq:q2k2withB} we obtain
\begin{align*}
q_2(\lambda) + &jk_2(\hat{\lambda}_L) \\
&= \hat{\lambda}_L \left( jh_2(0) - 2j\sin(\tfrac{\pi}{M}) \sum_{k=0}^{M-1} g(\psi_k) + o_P(1) \right) \\
&= \hat{\lambda}_L\big( jH + o_P(1) \big)
\end{align*}
and the lemma follows by taking real and imaginary parts.
\end{proof}

\begin{lemma}\label{lem:Blambdaconv}
With $B(\lambda)$ defined in~\eqref{eq:Blambdadefn} we have 
\[
B(\hat{\lambda}_L) = -\hat{\lambda}_L\left(2j\sin(\tfrac{\pi}{M}) \sum_{k=0}^{M-1} g(\psi_k) + o_P(1)\right) 
\]
\end{lemma}
\begin{proof}
Put $A(\lambda) = \expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)$.  Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$ we have
\begin{align*}
 A(\lambda)  &= \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) e^{j\fracpart{\phi}}\chi(\phi,\lambda) dr d\phi \\
&= \sum_{k\in\ints}\int_{0}^{2\pi} g(\phi) e^{j\fracpart{\phi}} \chi_k(\phi,\lambda) d\phi \\
&= \sum_{k=0}^{M-1}\int_{\psi_k-\lambda}^{\psi_k} g(\phi)  e^{j\fracpart{\phi}} d\phi,
\end{align*}
the last line because the $\chi_k(\phi,\lambda)$ terms inside the integral are zero for all $\phi \in [0,2\pi]$ when $k \notin \{0,\dots,M-1\}$.  Observe that $\fracpart{\phi} \rightarrow \tfrac{\pi}{M}$ as $\phi$ approaches $\psi_k$ from below.  Because $g(\psi_k)$ is continuous at $\psi_k$ for each $k = 0, \dots, M-1$ (by assumption in Theorem~\ref{thm:normality}) we have 
\[
\frac{1}{\lambda} A(\lambda) \rightarrow \sum_{k=0}^{M-1} g(\psi_k) e^{j\tfrac{\pi}{M}}
\]
as $\lambda$ approaches zero from above.  We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$ (see the proof of Lemma~\ref{lem:q2k2parts}).  The analogous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.  Thus
\[
A(\hat{\lambda}_L) = \hat{\lambda}_L \left(  e^{j\tfrac{\pi}{M}} \sum_{k=0}^{M-1} g(\psi_k) + o_P(1) \right)
\]
and the lemma follows since $\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) e^{j\frac{\pi}{M}} = -2j\sin(\tfrac{\pi}{M})$ and $B(\lambda) = \big( e^{-j\tfrac{2\pi}{M}} - 1 \big)A(\lambda)$.
\end{proof}

\begin{lemma}\label{lem:expectImRfracpart}
$\expect R_1 \sin\fracpart{\Phi_1} = 0$.
\end{lemma}
\begin{IEEEproof}
Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$ we have
\begin{align*}
\expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi.
\end{align*}
The proof is immediate since $g(\phi)$ is even and $\sin\fracpart{\phi}$ is odd.
% where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.  Proposition~\ref{prop:gg2cont} shows that $g$ is even on $(-\pi,\pi)$.  Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.  So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
% \[
% g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
% \]
% and
% \[
% \sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
% \]
% Now,
% \begin{align*}
% I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
% &\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
% &= 0
% \end{align*}
% for all integers $k$ because
% \begin{align*}
% \int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
% &= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
% \end{align*}
% Observe that the sets
% \[
%  [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
% \]
% partition $[-\pi, \pi]$, and so
% \[
% \expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
% \]
\end{IEEEproof}

@@

\begin{lemma}\label{lem:HLtoG} $\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1)$.
\end{lemma}
\begin{IEEEproof}
The argument is similar to that used in Lemma~\ref{lem:Qconv}.  First observe that
\begin{align*}
T_L(\lambda) &= \frac{\sabs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\sabs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
%&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
%& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
&=  \big( p + o(L^{-1/2}) \big) \expect R_1\cos(\lambda + \Phi_1) \\
&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1},
\end{align*}
and because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ (by assumption in Theorem~\ref{thm:normality}), we have
\begin{align*}
\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
\end{align*}
where
\begin{align}
q_1(\lambda) &= \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big) \;\;\; \text{and} \nonumber \\
q_2(\lambda) &= \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
\end{align}
Lemma~\ref{lem:q1k1parts} shows that $q_1(\hat{\lambda}_L) = \hat{\lambda}_Lo_P(1)$ and Lemma~\ref{lem:q2k2parts} shows that $q_2(\hat{\lambda}_L) = \hat{\lambda}_L o_P(1)$ and so
\[
\sqrt{L}\big( T_L(\lambda) - G(0) \big) =\sqrt{L}\hat{\lambda}_L o_P(1) + o(1).
\]
The lemma follows since $\sqrt{L}\hat{\lambda}_L$ converges in distribution.
\end{IEEEproof}
